Penalized linear mixed models for structured genetic data

### 2021-03-13 main.tex
Edits to tuning parameter selection, software

Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods, with many possible approaches. Here, for each model, we selected the $\lam$ value such that 50 SNPs were selected (recall that 50 is also the true number of causal SNPs). This method for selecting $\lambda$ was chosen in an attempt to compare the performance of all methods fairly, and as a strategy to evaluate methods in terms of variable selection, as it allows for direct comparison of the first 50 variables each method lets into the model.

Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons. Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods. \anna{ I want to cut the following sentence but I think we may need some other reason why we didn't do CV...} Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}. \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}

%------------------------------%
\subsection{\anna{Software}}
%------------------------------%
\label{sec:software}
\anna{Should this be here as a methods subsection or should it get its own section after discussion, before data availability?}

All analyses and simulations were performed using \texttt{R} version 3.6.3. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix was implemented using \texttt{ggmix} \citep{bhatnagar2020simultaneous}. The LMM-lasso-Rakitsch method was implemented using the \texttt{R} package, \texttt{penalizedLMM}, available on the github repository \url{https://github.com/areisett/penalizedLMM}. Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$. \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD.

\pb{BELOW IS JUST DUMP FROM EARLIER}
. This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package

A lasso model can then be fit using standard software for

A lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}.

LMM-lasso-ggmix iteratively updates $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm \citep{bhatnagar2020simultaneous}. This iterative update scheme prevents the use of standard lasso software, but is implemented in the \texttt{R} package \texttt{ggmix}. We further discuss differences between these methods in Section \ref{sec:results}.
\pb{END DUMP}

@@

\anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods. In a real data scenario, the true number of causal SNPs would not be known a priori. As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion. Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm. This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso. At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}

\anna{Several authors have used the Bayesian information criterion (BIC) for variable selection in mixed models \citep{bondell2010joint, schelldorfer2011estimation}. However it has been shown that the Akaike information criterion (AIC) and BIC may perform poorly in the high-dimensional setting \citep{fan2013tuning}. The high-dimensional BIC introduced by \citet{fan2013tuning} penalizes model parameters more severely in an attempt to perform better in high dimensional settings. \citet{bhatnagar2020simultaneous} employs this high-dimensional BIC as its default method to select $\lambda$.}

\anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice. \citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}. \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection. A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}

>>>>

Selection of the tuning parameter $\lambda$ plays an important role in all penalized regression methods.  Many approaches have been suggested in the literature, and can be roughly divided into four categories: (1) selecting a predetermined number of features, (2) using information criteria such as the Bayesian Information Criterion (BIC), (3) cross-validation and other resampling/sample-splitting approaches, and (4) inference-based approaches such as choosing the maximum $\lam$ that controls the false discovery rate (FDR) at a certain threshold.

These approaches have been extensively studied in the case of independent observations, but little work has been done on how they are affected by correlation among individuals.  We offer a few comments here, but note that our theoretical understanding remains relatively poor and there is no consensus on the optimal way to select $\lam$ in this case \citep{wauthier2013comparative}.

BIC has been used by several authors for variable selection in penalized models with correlated outcomes \citep{bondell2010joint, schelldorfer2011estimation}. BIC performs poorly in the high-dimensional setting, although modified high-dimensional variants have been proposed \citep{fan2013tuning,Chen2008a}. Cross-validation is also reasonable, with some caveats about the generalizability of prediction error \citep{Rabinowicz2020}. In the present context, this means that although cross-validation is a valid way to estimate prediction error of another sample from the same population, we cannot expect for example cross-validation of a GWAS from European samples to estimate prediction error accurately for calculating genetic risk in Chinese individuals.  Selecting $\lam$ on the basis of FDR is obviously attractive, although there has been little work extending these approaches to correlated outcomes.  Knockoff filters may be useful for this purpose \citep{Sesia2018}, although they have not been studied in the specific context of penalized LMMs.

Choosing a fixed number of variants, on the other hand, is straightforward and easy to apply to any model regardless of whether observations are correlated or not.  Its obvious drawback is that it requires an arbitrary decision regarding the number of variants to select, as opposed to the other, more data-driven methods.  However, these other methods often result in substantially different $\lam$ values being chosen from data set to data set or method to method, making comparisons difficult.  Fixing the number of variants selected provides an unambiguous way of comparing the accuracy of various methods.  Since comparing these methods is the goal of the present paper, we use this approach in all of our simulation studies, fixing the number of selected variants at 50 (the true number of causal SNPs).

%% Although cross-validation (CV) and information criterion-based methods are commonly used to select $\lambda$ values, in this context they have the potential to result in misleading comparisons. Because lasso, PC-lasso, and LMM-lasso use different likelihoods, an information-criterion-based approach such as BIC would not lend itself to comparable results across methods. \anna{ I want to cut the following sentence but I think we may need some other reason why we didn't do CV...} Meanwhile, whether cross-validation is advisable or not in the presence of dependent samples remains an open question \citep{roberts2017cross}. \anna{We include a more detailed discussion of tuning parameter selection methods for penalized LMMs in Section \ref{sec:discussion}.}

%% \anna{The nature of the simulations and analyses in this work have allowed us to select penalty parameters based on the true number of causal features in a model, which allows for fair comparison across methods. In a real data scenario, the true number of causal SNPs would not be known a priori. As such, appropriate methods for penalty parameter selection in penalized LMMs are worth additional discussion. Penalized LMMs are a special case of what is more broadly known in the statistical literature as a preconditioned lasso algorithm. This refers to the pre-multiplication of $\bX$ and $\by$ by some matrix prior to running the ordinary lasso. At present the theoretical understanding of these algorithms remains fairly poor \citep{wauthier2013comparative}, and there is no consensus on the optimal way to select a penalty parameter for such algorithms.}

%% \anna{Subsampling approaches, such as CV and stability selection are another common class of variable selection methods often preferred to information criterion-based approaches in practice. \citet{rakitsch2013lasso} used stability selection to evaluate the significance of individual SNPs \citep{meinshausen2010stability}. \citet{guo2019combining} used a combination of both methods for feature selection, first selecting several $\lambda$ values via CV, and then using these penalty parameters for stability selection and final feature selection. A rigorous assessment of penalty parameter selection in the context of penalized LMMs is beyond the scope of this work, but worthy of further investigation.}

%------------------------------%
\subsection{Software}
%------------------------------%
\label{sec:software}

All analyses and simulations were performed using R version 3.6.3. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix requires specialized software, but is implemented in the R package \texttt{ggmix} \citep{bhatnagar2020simultaneous}. Standard lasso software such as \texttt{glmnet} can be used to fit the LMM-lasso-Rakitsch method as well, although this requires a fair amount of preprocessing and setting up.  To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files. pb{Is this sentence important? Why do we care what parameterization Rakitsch used? Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD. \pb{Mention imputation capabilities?}

@@
